analysis

### 2020-06-05 analysis.Rmd
added floral taxonomy up to family level, tweaked some things in data_proc

### Prepare environment
```{r setup, include=FALSE}
knitr::opts_chunk$set(include = FALSE)
library(bipartite)
library(mgcv)
library(mgcViz)
#library(lme4)
#library(lmerTest)
library(tidyverse)
library(lubridate)
library(GGally)
```

@@

net <- read_csv("./processed_data/network.csv") %>%
  unite(site.year, site, year, sep = "_", remove = FALSE) %>% # create site.year field that can later be split back up
  unite(site.date, site, date, sep = "_", remove = FALSE) %>%
  mutate(year = factor(year)) %>% # convert year to factor
  full_join(bb_traits, by = c("bb.sp" = "bb.sp")) %>%
  left_join(site_data, by = "site") %>%
  left_join(gdd, by = c("site", "date")) %>%
  left_join(breaks, by = c("site", "year")) %>%
  ungroup() %>%
  mutate(phase = case_when(
    gdd.cum < break1.gdd ~ "founding",
    gdd.cum >= break1.gdd & gdd.cum < break2.gdd ~ "buildup",
    gdd.cum > break2.gdd ~ "reproductive"
  )) %>%
  mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive"))) %>%
  arrange(site, date)

survey <- read_csv("./processed_data/floral_survey.csv") %>%
  mutate(year = factor(year)) %>% # convert year to factor
  #semi_join(net, by = c("site", "plant.sp")) %>% # Consider only plants visited at a given site-date
  left_join(gdd, by = c("site", "date")) %>%
  left_join(breaks, by = c("site", "year")) %>%
  left_join(site_data, by = "site") %>%
  ungroup() %>%
  mutate(phase = case_when(
    gdd.cum < break1.gdd ~ "founding",
    gdd.cum >= break1.gdd & gdd.cum < break2.gdd ~ "buildup",
    gdd.cum > break2.gdd ~ "reproductive"
  )) %>%
  mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive"))) %>%
  arrange(site, date)
```
### BB and floral turnover ~ elevation
```{r, include=FALSE}
bb_range <- net %>%
  group_by(site, elev.mean, bb.sp) %>%
  summarize(abund = n()) %>%
  group_by(bb.sp) %>%
  mutate(elev.floor = min(elev.mean),
         elev.ceiling = max(elev.mean),
         elev.range = elev.ceiling - elev.floor,
         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2))

fl_survey_elev <- survey %>%
  filter(flower.cover > 0) %>%
  group_by(site, plant.sp) %>%
  summarize(abund = sum(flower.cover)) %>%
  left_join(site_data) %>%
  semi_join(net, by = "plant.sp") %>% # constrain to plant species that were actually visited
  group_by(plant.sp) %>%
  mutate(elev.floor = min(elev.mean),
         elev.ceiling = max(elev.mean),
         elev.range = elev.ceiling - elev.floor,
         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2),
         sites = length(unique(site)))

@@

@@

>>>>

### Prepare environment
```{r setup, include=FALSE}
knitr::opts_chunk$set(include = FALSE)
library(bipartite)
library(mgcv)
library(mgcViz)
#library(lme4)
#library(lmerTest)
install.packages("taxize")
library(taxize)
library(tidyverse)
library(lubridate)
library(GGally)
```

@@

fl_traits <- read.csv("./processed_data/floral_k_type.csv") %>%
  select(plant.sp, k.type)

fl_tax <- read_csv("./processed_data/floral_tax.csv")

net <- read_csv("./processed_data/network.csv") %>%
  unite(site.year, site, year, sep = "_", remove = FALSE) %>% # create site.year field that can later be split back up
  unite(site.date, site, date, sep = "_", remove = FALSE) %>%
  mutate(year = factor(year)) %>% # convert year to factor
  full_join(bb_traits, by = c("bb.sp" = "bb.sp")) %>%
  left_join(site_data, by = "site") %>%
  left_join(gdd, by = c("site", "date")) %>%
  left_join(breaks, by = c("site", "year")) %>%
  left_join(fl_traits) %>%
  left_join(fl_tax, by = "plant.genus") %>%
  ungroup() %>%
  mutate(phase = case_when(
    gdd.cum < break1.gdd ~ "founding",
    gdd.cum >= break1.gdd & gdd.cum < break2.gdd ~ "buildup",
    gdd.cum > break2.gdd ~ "reproductive"
  )) %>%
  mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive"))) %>%
  arrange(site, date)

survey <- read_csv("./processed_data/floral_survey.csv") %>%
  mutate(year = factor(year)) %>% # convert year to factor
  semi_join(net, by = "plant.sp") %>% # Consider only plants visited at a given site-date
  left_join(gdd, by = c("site", "date")) %>%
  left_join(breaks, by = c("site", "year")) %>%
  left_join(site_data, by = "site") %>%
  left_join(fl_traits) %>%
  left_join(fl_tax) %>%
  ungroup() %>%
  mutate(phase = case_when(
    gdd.cum < break1.gdd ~ "founding",
    gdd.cum >= break1.gdd & gdd.cum < break2.gdd ~ "buildup",
    gdd.cum > break2.gdd ~ "reproductive"
  )) %>%
  mutate(phase = factor(phase, levels = c("founding", "buildup", "reproductive"))) %>%
  arrange(site, date)

```
### BB and floral turnover ~ elevation
```{r, include=FALSE}
bb_range <- net %>%
  group_by(site, elev.mean, bb.sp) %>%
  summarize(abund = n()) %>%
  group_by(bb.sp) %>%
  mutate(elev.floor = min(elev.mean),
         elev.ceiling = max(elev.mean),
         elev.range = elev.ceiling - elev.floor,
         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2))

fl_survey_elev <- survey %>%
  filter(flower.cover > 0) %>%
  group_by(site, plant.sp) %>%
  summarize(abund = sum(flower.cover)) %>%
  left_join(site_data) %>%
  group_by(plant.sp) %>%
  mutate(elev.floor = min(elev.mean),
         elev.ceiling = max(elev.mean),
         elev.range = elev.ceiling - elev.floor,
         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2),
         sites = length(unique(site)))

fl_survey_elev_genus <- survey %>%
  filter(flower.cover > 0) %>%
  group_by(site, genus) %>%
  summarize(abund = sum(flower.cover)) %>%
  left_join(site_data) %>%
  group_by(genus) %>%
  mutate(elev.floor = min(elev.mean),
         elev.ceiling = max(elev.mean),
         elev.range = elev.ceiling - elev.floor,
         elev.med = elev.floor + ((elev.ceiling - elev.floor)/2),
         sites = length(unique(site)))

@@

ggplot(filter(fl_survey_elev_genus, sites > 0), 
       aes(reorder(genus, elev.med), elev.mean)) +
  geom_line(size = 1, color = "gray60") +
  geom_point(alpha = 0.5) +
  theme_light(10) +
  #theme(axis.text.x = element_blank()) +
  xlab("Plant genus") +
  ylab("Elevation") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

@@

### FL traits by elevation -- do we see a sharp change at the tree line?
```{r}
survey_trait_elev <- survey %>%
  group_by(site, year, elev.mean,  k.type) %>%
  summarize(abund = sum(flower.cover)) %>%
  group_by(site, elev.mean, year) %>%
  mutate(div = length(unique(k.type)))

net_trait_elev <- net %>%
  group_by(site, year, elev.mean,  k.type) %>%
  summarize(abund = n()) %>%
  group_by(site, elev.mean, year) %>%
  mutate(div = length(unique(k.type)))

survey_genus_elev <- survey %>%
  group_by(site, year, elev.mean, plant.genus) %>%
  summarize(abund = sum(flower.cover)) %>%
  group_by(site, elev.mean, year) %>%
  mutate(div = length(unique(plant.genus)))

survey_family_elev <- survey %>%
  group_by(site, year, elev.mean,  plant.family) %>%
  summarize(abund = sum(flower.cover)) %>%
  group_by(site, elev.mean, year) %>%
  mutate(div = length(unique(plant.family)))

net_genus_elev <- survey %>%
  group_by(site, year, elev.mean,  plant.genus) %>%
  summarize(abund = n()) %>%
  group_by(site, elev.mean, year) %>%
  mutate(div = length(unique(plant.genus)))

net_family_elev <- survey %>%
  group_by(site, year, elev.mean,  plant.family) %>%
  summarize(abund = n()) %>%
  group_by(site, elev.mean, year) %>%
  mutate(div = length(unique(plant.family)))

ggplot(survey_trait_elev, aes(elev.mean, log(abund))) +
  geom_point() +
  stat_smooth() +
  facet_wrap(~k.type, scales = "free_y")

ggplot(net_trait_elev, aes(elev.mean, log(abund))) +
  geom_point() +
  stat_smooth() +
  facet_wrap(~k.type, scales = "free_y")

ggplot(survey_trait_elev, aes(elev.mean, div)) +
  geom_point() +
  stat_smooth()

ggplot(net_trait_elev, aes(elev.mean, div)) +
  geom_point() +
  stat_smooth()

ggplot(survey_genus_elev, aes(elev.mean, log(abund))) +
  geom_point() +
  stat_smooth() +
  facet_wrap(~plant.genus, scales = "free_y")

ggplot(survey_family_elev, aes(elev.mean, log(abund))) +
  geom_point() +
  stat_smooth() +
  facet_wrap(~plant.family, scales = "free_y")

ggplot(survey_genus_elev, aes(elev.mean, div)) +
  geom_point() +
  stat_smooth()

ggplot(survey_family_elev, aes(elev.mean, div)) +
  geom_point() +
  stat_smooth()

ggplot(net_genus_elev, aes(elev.mean, log(abund))) +
  geom_point() +
  stat_smooth() +
  facet_wrap(~plant.genus, scales = "free_y")

ggplot(net_family_elev, aes(elev.mean, log(abund))) +
  geom_point() +
  stat_smooth() +
  facet_wrap(~plant.family, scales = "free_y")

ggplot(net_genus_elev, aes(elev.mean, log(abund))) +
  geom_point() +
  stat_smooth(se = FALSE) +
  facet_wrap(~plant.genus)

ggplot(net_genus_elev, aes(elev.mean, log(abund))) +
  geom_point() +
  stat_smooth() +
  facet_wrap(~plant.genus, scales = "free_y")
ggplot(net_genus_elev, aes(elev.mean, abund)) +
  geom_point() +
  stat_smooth(se = FALSE) +
  facet_wrap(~plant.genus)

ggplot(net_genus_elev, aes(elev.mean, div)) +
  geom_point() +
  stat_smooth()
```
